---
title: "STA141B-HW4"
author: "Stella Liu"
date: "2026-05-26"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results='asis', echo=FALSE, warning = FALSE, message = FALSE}
library(httr2)
library(jsonlite)
library(knitr)
```

Assignment 4: 

1. Get the information in the “table” on this page, i.e., the data for the Likely Democratic, Competitive States, Likely Republican for the presidential election.

```{r, results='asis', echo=FALSE, warning = FALSE, message = FALSE}
#retrieve response from web link network while stimulating standard browser interface
pre_url <- "https://apps.npr.org/2024-election-results/data/president.json"
req <- request(pre_url) %>% req_user_agent("Mozilla/5.0")
resp <- req_perform(req)

#binary network signal into active nested R list text strings 
master_list <- fromJSON(resp_body_string(resp))
df_results <- master_list$results
names(df_results)
unique(df_results$rating)
```

```{r, results='asis', echo=FALSE, warning = FALSE, message = FALSE}
likely_dem <- subset(df_results, rating %in% c("likely-d", "lean-d"))
competitive <- subset(df_results, rating == "toss-up")
likely_rep <- subset(df_results, rating %in% c("likely-r", "lean-r"))

print(likely_dem[, c("stateName", "rating")])
print(competitive[, c("stateName", "rating")])
print(likely_rep[, c("stateName", "rating")])
```

2. For each state,
• for the presidential election,
– get the “Demographics by County” information,
– create the table for the overall result for the state.

```{r, results='asis', echo=FALSE, warning = FALSE, message = FALSE}
df_results_pres <- fromJSON(resp_body_string(resp))$results
distinct_states <- df_results_pres[!duplicated(df_results_pres$state), ]

#state by state cycle for presidential 
for (i in 1:nrow(distinct_states)) {
  #identification flags
  st_name <- distinct_states$stateName[i]
  st_code <- toupper(df_results_pres$state[i])
  
  if (st_code == "US" || st_code == "") next
  
  cat("\n## ", st_name, " Presidential Results\n")
  
  state_overall_votes <- df_results_pres$candidates[[i]]
  cat("#### Overall State Totals\n")
  
  if (!is.null(state_overall_votes) && "votes" %in% names(state_overall_votes)) {
    total_state_votes <- sum(as.numeric(state_overall_votes$votes), na.rm = TRUE)
    
    if (total_state_votes > 0) {
      #calculate percentage for each
      state_overall_votes$percentage <- paste0(
        round((as.numeric(state_overall_votes$votes) / total_state_votes) * 100, 1), "%"
      )
    } else {
      state_overall_votes$percentage <- "0.0%"
    }
    
    print(kable(head(state_overall_votes[, c("last", "party", "votes", "percentage")], 5), row.names = FALSE))
  }

#overall validation
if (!is.null(state_overall_votes) && "votes" %in% names(state_overall_votes)) {
  
  total_votes <- sum(as.numeric(state_overall_votes$votes), na.rm = TRUE)
  percents <- as.numeric(gsub("%", "", state_overall_votes$percentage))
  
  cat("**Overall Validation Checks:**\n")
  
  cat("- Percent sum:", round(sum(percents, na.rm = TRUE), 1), "%\n")
  
  cat("- Total Overall Votes:", total_votes, "\n\n")
}
  
  #use state code for specialized URL
  state_url <- paste0("https://apps.npr.org/2024-election-results/data/counties/", st_code, "-0.json")
  
  #protective: if missing return null
  state_data <- tryCatch({
    fromJSON(resp_body_string(req_perform(req_user_agent(request(state_url), "Mozilla/5.0"))))
  }, error = function(e) { NULL })
  
  #verify data integrity 
  raw_df <- if (!is.null(state_data)) state_data$results else NULL
  
  #filter rows with indicator p for presidential, track county
  if (!is.null(raw_df) && is.data.frame(raw_df)) {
    sub_df <- raw_df[trimws(tolower(raw_df$office)) == "p" & trimws(tolower(raw_df$level)) == "county", ]
    if (nrow(sub_df) == 0) sub_df <- raw_df[raw_df$level == "county" | !is.null(raw_df$reportingunitName), ]
    
    #table of demographic variables
    if (nrow(sub_df) > 0 && !is.null(sub_df$county)) {
      clean_demographics <- data.frame(
        County       = if(!is.null(sub_df$county$countyName)) sub_df$county$countyName else sub_df$reportingunitName,
        Median_Income       = if(!is.null(sub_df$county$median_income)) sub_df$county$median_income else NA,
        Black_Pop    = if(!is.null(sub_df$county$black_total)) sub_df$county$black_total else NA,
        Hispanic_Pop = if(!is.null(sub_df$county$hispanic_total)) sub_df$county$hispanic_total else NA,
        Votes        = if(!is.null(sub_df$totalVotes)) sub_df$totalVotes else NA,
        stringsAsFactors = FALSE
      )
      #remove missing 
      clean_demographics <- clean_demographics[!is.na(clean_demographics$County), ]
      
      cat("\n#### County Demographics\n")
      print(kable(head(clean_demographics, 5), row.names = FALSE))
      
#county validation
if (!is.null(clean_demographics) && nrow(clean_demographics) > 0) {
  
  cat("**County Validation Checks:**\n")
  
  cat("- Total County Votes:", sum(clean_demographics$Votes, na.rm = TRUE), "\n")
  
  cat("\n")
}
    #protective: if missing return message 
    } else {
      cat("\n*Demographic columns missing inside this file structure.*\n")
    }
  } else {
    cat("\n*Network dataset could not be pulled.*\n")
  }
  cat("\n---\n")
}
```

• same for the Senate elections.
– “Demographic by County” information
– overall result for the states

```{r, results='asis', echo=FALSE, warning=FALSE, message=FALSE}
sen_url <- "https://apps.npr.org/2024-election-results/data/senate.json"
req1 <- request(sen_url) %>% req_user_agent("Mozilla/5.0")
resp1 <- req_perform(req1)
master_list1 <- fromJSON(resp_body_string(resp1))

#check structural variation 
if ("races" %in% names(master_list1)) {
  df_results1 <- master_list1$races
} else if ("results" %in% names(master_list1)) {
  df_results1 <- master_list1$results
} else {
  df_results1 <- master_list1
}

#protective: if missing return message
if (!is.data.frame(df_results1) || nrow(df_results1) == 0) {
  cat("### No race data found.\n")
  knitr::knit_exit()
}

#state by state cycle for senate 
for (i in seq_len(nrow(df_results1))) {
  race <- df_results1[i, ]

  st_name <- if (!is.null(race$stateName) && !is.na(race$stateName)) race$stateName else race$state

  cat(paste0("\n## ", st_name, " Senate Results\n"))

  #candidate count
  cands <- race$candidates
  if (is.list(cands) && length(cands) > 0) {
    if (!is.data.frame(cands)) cands <- as.data.frame(do.call(rbind, cands))
    
    if ("votes" %in% names(cands)) {
      total_senate_votes <- sum(as.numeric(cands$votes), na.rm = TRUE)
      
      if (total_senate_votes > 0) {
        #calculate percentage for each
        cands$percentage <- paste0(
          round((as.numeric(cands$votes) / total_senate_votes) * 100, 1), "%"
        )
      } else {
        cands$percentage <- "0.0%"
      }
    }
    
    valid_cols <- intersect(names(cands), c("last", "party", "votes", "percentage", "winner"))
    
    if (length(valid_cols) > 0) {
      cat("#### Overall State Totals\n")
      print(kable(head(cands[, valid_cols, drop = FALSE], 6),
                  row.names = FALSE, format = "markdown"))
    }
  }

#overall validation
if (!is.null(cands) && "votes" %in% names(cands)) {
  
  total_votes <- sum(as.numeric(cands$votes), na.rm = TRUE)
  percents <- as.numeric(gsub("%", "", cands$percentage))
  
  cat("**Overall Validation Checks:**\n")
  
  cat("- Percent sum:", round(sum(percents, na.rm = TRUE), 1), "%\n")
  
  cat("- Total votes:", total_votes, "\n\n")
}

  #identification flags
  st_code_sen <- toupper(race$state)
  race_id     <- race$id

  #use state code and race id for specialized URL
  state_url_sen <- paste0(
    "https://apps.npr.org/2024-election-results/data/counties/",
    st_code_sen, "-", race_id, ".json"
  )
  
  #protective: if missing return null
  state_data <- tryCatch({
    req_sen <- request(state_url_sen) %>% req_user_agent("Mozilla/5.0")
    fromJSON(resp_body_string(req_perform(req_sen)))
  }, error = function(e) NULL)

  raw_df_sen <- if (!is.null(state_data)) state_data$results else NULL

  #both country and township level tracking 
  if (!is.null(raw_df_sen) && is.data.frame(raw_df_sen)) {
    sub_df_sen <- raw_df_sen[!trimws(tolower(raw_df_sen$level)) %in% c("state", "national", "us"), ]
    
    #table of demographic variables 
    if (nrow(sub_df_sen) > 0 && "county" %in% names(sub_df_sen) && !is.null(sub_df_sen$county)) {
      clean_demographics_sen <- data.frame(
        'County/Township' = if (!is.null(sub_df_sen$county$countyName)) sub_df_sen$county$countyName else sub_df_sen$reportingunitName,
        Income       = if (!is.null(sub_df_sen$county$median_income)) sub_df_sen$county$median_income else NA,
        Black_Pop    = if (!is.null(sub_df_sen$county$black_total)) sub_df_sen$county$black_total else NA,
        Hispanic_Pop = if (!is.null(sub_df_sen$county$hispanic_total)) sub_df_sen$county$hispanic_total else NA,
        Votes        = if (!is.null(sub_df_sen$totalVotes)) sub_df_sen$totalVotes else NA,
        stringsAsFactors = FALSE
      )

      #protective: if missing return message 
      if (nrow(clean_demographics_sen) > 0) {
        cat("\n#### County/Township Demographics\n")
        print(kable(head(clean_demographics_sen, 5),
                    row.names = FALSE, format = "markdown"))
        
#county validation
if (!is.null(clean_demographics_sen) && nrow(clean_demographics_sen) > 0) {
  
  cat("**County Validation Checks:**\n")
  
  cat("- Total County Votes:", sum(clean_demographics_sen$Votes, na.rm = TRUE), "\n")
  
  cat("\n")
}
        
      } else {
        cat("\n*No demographic available for this state.*\n")
      }
    } else {
      cat("\n*Demographic missing for this state.*\n")
    }
  } else {
    cat(paste0("\n*Failed to parse for ", st_name, ".*\n"))
  }

  cat("\n---\n")
}
```

I used Gemini, chat link: https://gemini.google.com/app/933ebbb21f235908

I input verbal descriptions of what I want to achieve and ask it for brainstorming ideas and approaches. I check the code output by trying it and visualizing output segment by segment, I also ask it to explain the output code line by line so I'm not blindly copy pasting, but learning the code and understanding each line and function's purpose. 
There were many redundancies in code that was given as feedback, I also made changes to the code to better fit the structure of the data. I also often edited and reformatted for personal consistency and preference, removing identical library blocks and data blocks. 
Some errors I've encountered and fixed, I first failed to take into consideration that for senate results, some states had township instead of county, which was later updated with a new sub_df_sen line. I later decided to include the winner and percentage for each overall state total tables, this way we can also validate that there is one winner among all the candidates, and that the percentage adds up to 100%. I also validated by double checking between sites for accurate information and representation.
I worked on this assignment on and off throughout the weeks, intensifying to around 5 hours everyday since the Friday before the deadline.

```{r, results='asis', echo=FALSE}
cat("\\newpage") #new page
```

# Appendix: R code

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```